---
title: "Data for dissimilarity calculations"
author: "Jon Minton"
date: "16 June 2015"
output: html_document
---

# Introduction
A number of datasets have been identified that can be used to calculate dissimilarity scores using. These data sets are derived mainly from the 2001 and 2011 censuses, and the Scottish Neighbourhood Statistics (SNS) website. Where possible, I have used data relating to 2001 and 2011 for comparability with census variables. 
The tables have been prepared mainly for the 2196 datazones that comparise a definition of Greater Glasgow based on health boards. However in principle all tables could be prepared for the whole of Scotland.


# Census-based variables: country of origin, ethnic group, and religion

## Country of origin, ethnic group and religion.

These tables were first found and used as part of the dissimilarity inference app, and are available 
to select within the latest version of the Shiny app. 


## Country of origin
```{r}
require(readr)
require(dplyr)
require(tidyr)
dta_coo <- read_csv("data/derived/coo.csv")
dta_coo
dta_coo %>% group_by(year) %>% tally
dta_coo %>% select(-1) %>% gather(key="coo", value="count", -datazone, -year) %>% 
    group_by(year, coo) %>% summarise(count=sum(count))

```

## The ethnic group table:

```{r}
dta_eg <- read_csv("data/derived/ethnicity.csv")
dta_eg %>% group_by(year) %>% tally
dta_eg %>% gather(key= "eg", value="count", -datazone, -year) %>% xtabs(count ~ year + eg, data= . )

```

## Religion table:

```{r}
dta_rel <- read_csv("data/derived/rel.csv")
dta_rel
dta_rel %>% gather(key= "rel", value="count", -datazone, -year) %>% xtabs(count ~ year + rel, data= . )
```


# Additional census variables

These variables were identified more recently. 


## Highest qualification
 
```{r}
 
 dta_qual <- read_csv("data/derived/highest_qual.csv")
dta_qual
dta_qual %>% gather(key="hq", value="count", -datazone, -year) %>% xtabs(count ~ year + hq, data = .)
 
```

## Industry

```{r}
dta_ind <- read_csv("data/derived/industry.csv")
dta_ind
dta_ind %>% gather(key="industry", value="count", -datazone, -year) %>% xtabs(count ~ industry + year, data = .)
```


## Economic activity

```{r}
dta_econ <- read_csv("data/derived/economic_activity.csv")
dta_econ

dta_econ %>% gather(key="econ", value="count", -datazone, -year) %>% xtabs(count ~ econ + year, data = .)
```

## Socioeconomic class 

```{r}
dta_sec <- read_csv("data/derived/sec_by_dz.csv")
dta_sec
dta_sec %>% gather(key="sec", value = "count", -datazone, -year) %>% xtabs(count ~ year + sec, data = .)

```

# Tables from the SNS and elsewhere 

A number of tables refer to houses rather than individuals or households. These are mainly taken from the SNS website.

## Dwellings by size (number of rooms) - the earliest available year is 2006, unfortunately.
```{r}
dta_dwelsize <- read_csv("data/derived/dwellings_by_size.csv")
dta_dwelsize
dta_dwelsize %>% xtabs(count ~ num_of_rooms + year, data = .)
```

## Dwellings by council tax band

```{r}
dta_band <- read_csv("data/derived/dwellings_by_band.csv")
dta_band
dta_band %>% gather(key="band", value="count", -datazone, -year) %>% xtabs(count ~ year + band, data = .)

```

## Dwellings by type

```{r}
dta_type <- read_csv("data/derived/dwellings_by_type.csv")
dta_type
dta_type %>% xtabs(count ~ year + type, data = .)
```




# Demographic variables

I have data on population, by five year age groups and gender, at datazone level, based on a combination of the 2001 and 2011 censuses, and small area population estimates (SAPE) for non-census years. I have combined into a single, very large file, as well as used this file to group people into different categories based on age and gender classifications. 

```{r}
demo_big <- read_csv("data/derived/populations_by_age_year_sex.csv")
demo_big
demo_big %>% xtabs(count ~ year + age_range + sex, data = .)
```

As you can see, the age groupings are net entirely consistent across years. I have produced a version with some consistent groupings, just for 2001 and 2011, here

```{r}
demo_grps <- read_csv("data/derived/demographic_groupings.csv")
demo_grps

demo_grps %>% gather(key="grp", value="count", -dz_2001, -year) %>% xtabs(count ~ year + grp, data = .)

# The code used to produce the groupings:
#         f1 = female_0_4 + female_5_9 + female_10_14 + female_10_15,
#         f2 = female_15_19 + female_16_19 + female_20_24,
#         f3 = female_25_29 + female_30_34 + female_35_39,
#         f4 = female_40_44 + female_45_49 + female_50_54,
#         f5 = female_55_59 + female_60_64,
#         f6 = female_65_69 + female_70_74 + female_75_79,
#         f7 = female_80_84 + female_85_89 + female_90_101,
#         
#         m1 = male_0_4 + male_5_9 + male_10_14 + male_10_15,
#         m2 = male_15_19 + male_16_19 + male_20_24,
#         m3 = male_25_29 + male_30_34 + male_35_39,
#         m4 = male_40_44 + male_45_49 + male_50_54,
#         m5 = male_55_59 + male_60_64,
#         m6 = male_65_69 + male_70_74 + male_75_79,
#         m7 = male_80_84 + male_85_89 + male_90_101
```

# Quirkier tables

I have also extracted data from the Postcode Address File (PAF) on different types of building use.
The PAF distinguishes between 

* *address points*: residential
* *delivery points*: any use
* *small business points*: small businesses

The number of delivery points less the number of address points and small business points should 
indicate the number of large businesses. But in practice (I *think*) categories are not mutually exclusive,
so correctly using the PAF can be tricky...

```{r}

bld_use <- read_csv("data/derived/building_use.csv")
bld_use 
bld_use %>% gather(key="type", value="count", -datazone, -year) %>% xtabs(count ~ year + type, data = . )
```


#Other variables

A very large number of variables have been extracted and rearranged from the SNS. These are available on Dropbox. 

