---
title: "diversity means and sds"
author: "Jon Minton"
date: "Tuesday, March 10, 2015"
output: word_document
---

# Request from Ade

Jon,
 
Meantime, do you think you could also produce a table showing the average (mean) values and s.d.s for the diversity scores on each dimension, for  all datazones in Greater Glasgow  and then also separately by SIMD 2012 quintile?
 
Yours,
Ade

#Mean and SDs for diversity scores

```{r prereqs, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# What are the sources of data?
rm(list=ls())


setwd("E:/repos/gowell")

require(knitr)
require(repmis)
require(plyr)
require(tidyr)
require(stringr)
require(dplyr)

require(ggplot2)
require(rgl)

require(vegan)

demo_div <- read.csv("data/derived/demographic_diversity_2010.csv") %>%
    tbl_df %>%
    rename(demographic=diversity) %>%
    select(datazone, demographic)


dtype_div <- read.csv("data/derived/diversity_dwelling_type_2011.csv") %>%
    tbl_df %>%
    rename(dtype=diversity) %>%
    select(datazone, dtype)

eth_div <- read.csv("data/derived/diversity_ethnicity_2011.csv") %>%
    tbl_df %>%
    rename(ethnicity=diversity) %>%
    select(datazone, ethnicity)

sec_div <- read.csv("data/derived/diversity_sec_by_dz_2011_census.csv") %>%
    tbl_df %>%
    rename(sec=sec_div) %>%
    select(datazone, sec)

ten_div <- read.csv("data/derived/diversity_tenure_2001.csv" ) %>%
    tbl_df %>%
    rename(tenure=diversity) %>%
    select(datazone, tenure)

spac_div <- read.csv("data/derived/diversity_space_2001.csv") %>%
    tbl_df %>%
    rename(space=diversity) %>%
    select(datazone, space)

band_div <- read.csv("data/derived/diversity_dwelling_band_2011.csv") %>%
    tbl_df %>%
    rename(band=diversity) %>%
    select(datazone, band)

joined <- Reduce(inner_join, list(demo_div, dtype_div, eth_div, sec_div, ten_div, spac_div, band_div))
```

##Mean and SD of raw results

The following shows the means and sds for each diversity type. The first table 
shows the raw scores; the second shows each after being converted to a percentile rank.
The third shows the scores after being converted to the proportion of the range 
from the minimum to maximum values (i.e. x - min(x) / (max(x) -min(x)))

```{r table_raw, echo=FALSE, results="asis"}


setwd("E:/repos/gowell")

summaries_raw <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 

summaries_normalised <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>% 
    mutate_each(funs(percent_rank), -datazone) %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 

fn <- function(x){
    out <- (x-min(x))/(max(x)-min(x))
    return(out)
}

summaries_normalised2 <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>% 
    mutate_each(funs(fn), -datazone) %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 

kable(summaries_raw, type="markdown", digits=3)
kable(summaries_normalised, type="markdown", digits=3)
kable(summaries_normalised2, type="markdown", digits=3)

```


##Results by SIMD quintile

Here are the diversity scores by SIMD quintile. The first plot presents the 'raw' results. 
In the second plot all values are presented as rank percentiles (i.e. 1 is highest rank, 0 lowest rank).
In the third plot all values are converted to proportions using (x-min(x))/(max(x)-min(x))
```{r table_quintile, echo=FALSE, results="asis"}

setwd("E:/repos/gowell")
simd_raw <- read.csv("data/simd/00410767.csv", header=T) %>%
    tbl_df

simd <- simd_raw %>%
    select(datazone=Data.Zone, la=Local.Authority.Name, simd=Overall.SIMD.2012.Score)


greater_glasgow_dzs <- read.csv("data/geographies/dzs_in_greater_glasgow.csv")  %>% tbl_df %>%
    rename(datazone=dz_2001)

simd <- simd %>%
    right_join(greater_glasgow_dzs) %>%
    select(-chp)

simd <- simd  %>% mutate(simd_q=ntile(simd, 5))

joined <- joined  %>%  inner_join(simd)

summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value)) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd, yend=mean+2*sd)) +
    facet_wrap(~ type)

summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    mutate_each(vars=c(-datazone, -simd_q), funs(percent_rank)) %>%
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value), n=n()) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd/sqrt(n), yend=mean+2*sd/sqrt(n))) +
    facet_wrap(~ type)

fn <- function(x){
    out <- (x-min(x))/(max(x)-min(x))
    return(out)
}
summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    mutate_each(vars=c(-datazone, -simd_q), funs(fn)) %>%
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value), n=n()) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd/sqrt(n), yend=mean+2*sd/sqrt(n))) +
    facet_wrap(~ type)

```

# Entropy scores

As above, but using entropy scores 

```{r entropy, cache=TRUE, message=FALSE, echo=FALSE}

# What are the sources of data?
rm(list=ls())

setwd("E:/repos/gowell")

require(knitr)
require(repmis)
require(plyr)
require(tidyr)
require(stringr)
require(dplyr)

require(ggplot2)
require(rgl)

require(vegan)

demo_div <- read.csv("data/derived/demographic_diversity_2010.csv") %>%
    tbl_df %>%
    rename(demographic=H) %>%
    select(datazone, demographic)


dtype_div <- read.csv("data/derived/diversity_dwelling_type_2011.csv") %>%
    tbl_df %>%
    rename(dtype=H) %>%
    select(datazone, dtype)

eth_div <- read.csv("data/derived/diversity_ethnicity_2011.csv") %>%
    tbl_df %>%
    rename(ethnicity=H) %>%
    select(datazone, ethnicity)

sec_div <- read.csv("data/derived/diversity_sec_by_dz_2011_census.csv") %>%
    tbl_df %>%
    rename(sec=sec_h) %>%
    select(datazone, sec)

ten_div <- read.csv("data/derived/diversity_tenure_2001.csv" ) %>%
    tbl_df %>%
    rename(tenure=H) %>%
    select(datazone, tenure)

spac_div <- read.csv("data/derived/diversity_space_2001.csv") %>%
    tbl_df %>%
    rename(space=H) %>%
    select(datazone, space)

band_div <- read.csv("data/derived/diversity_dwelling_band_2011.csv") %>%
    tbl_df %>%
    rename(band=H) %>%
    select(datazone, band)

joined <- Reduce(inner_join, list(demo_div, dtype_div, eth_div, sec_div, ten_div, spac_div, band_div))
```


## Mean and SD (Entropy scores)

##Mean and SD of raw results

The following shows the means and sds for each diversity type. The first table 
shows the raw scores; the second shows each after being converted to a percentile rank. 

```{r table_entropy, echo=FALSE, results="asis"}


setwd("E:/repos/gowell")

summaries_raw <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 

summaries_normalised <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>% 
    mutate_each(funs(percent_rank), -datazone) %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 

summaries_normalised2 <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band)  %>% 
    mutate_each(funs(fn), -datazone) %>%
    gather(key=type, value=value, -datazone) %>%
    group_by(type) %>%
    summarise(mean=mean(value), sd=sd(value)) 


kable(summaries_raw, type="markdown", digits=3)
kable(summaries_normalised, type="markdown", digits=3)
kable(summaries_normalised2, type="markdown", digits=3)

```


##Results by SIMD quintile

Here are the diversity scores by SIMD quintile. The first plot presents the 'raw' results. 
In the second plot all values are presented as rank percentiles (i.e. 1 is highest rank, 0 lowest rank).
In the third plot all values are converted to proportions using (x-min(x))/(max(x)-min(x))
```{r table_quintile_entropy, echo=FALSE, results="asis"}

setwd("E:/repos/gowell")
simd_raw <- read.csv("data/simd/00410767.csv", header=T) %>%
    tbl_df

simd <- simd_raw %>%
    select(datazone=Data.Zone, la=Local.Authority.Name, simd=Overall.SIMD.2012.Score)


greater_glasgow_dzs <- read.csv("data/geographies/dzs_in_greater_glasgow.csv")  %>% tbl_df %>%
    rename(datazone=dz_2001)

simd <- simd %>%
    right_join(greater_glasgow_dzs) %>%
    select(-chp)

simd <- simd  %>% mutate(simd_q=ntile(simd, 5))

joined <- joined  %>%  inner_join(simd)

summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value)) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd, yend=mean+2*sd)) +
    facet_wrap(~ type)

summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    mutate_each(vars=c(-datazone, -simd_q), funs(percent_rank)) %>%
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value), n=n()) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd/sqrt(n), yend=mean+2*sd/sqrt(n))) +
    facet_wrap(~ type)


fn <- function(x){
    out <- (x-min(x))/(max(x)-min(x))
    return(out)
}
summaries_quints <- joined  %>% 
    select(datazone, demographic, dtype, ethnicity, 
           sec, tenure, space, band, simd_q)  %>% 
    mutate_each(vars=c(-datazone, -simd_q), funs(fn)) %>%
    gather(key=type, value=value, -datazone, -simd_q) %>%
    group_by(type, simd_q) %>%
    summarise(mean=mean(value), sd=sd(value), n=n()) 

summaries_quints %>%
    ggplot +
    geom_line(aes(x=simd_q, y=mean)) +
    geom_segment(aes(x=simd_q, xend=simd_q, y=mean-2*sd/sqrt(n), yend=mean+2*sd/sqrt(n))) +
    facet_wrap(~ type)
```